---
title: "Reproducible Research: Peer Assessment 1"
author: "Kpakpo Moevi"
date: "June 6, 2017"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
#1.Code for reading in the dataset and/or processing the data
Load the data 
```{r , echo = TRUE}
if (!file.exists("activity.csv")) {
  activity<-  unzip("activity.zip")
}
activity_moni_data <- read.csv("activity.csv")

```
Process the data
```{r ,echo =TRUE}
 activity_moni_data$date<-as.Date(activity_moni_data$date)
 activity_moni_data$interval<-as.factor(activity_moni_data$interval)
 head(activity_moni_data$date)
 head(activity_moni_data$interval)
```
 Load package plyr


```{r, echo = TRUE}
library(plyr)
number_of_steps_per_day<-ddply(activity_moni_data,.(date),summarise,sum = sum(steps,na.rm = TRUE))
head(number_of_steps_per_day)
#2.Histogram of the total number of steps taken each day

``` {r, echo =TRUE}

hist(number_of_steps_per_day$sum,col = "blue",main = "Total number of steps taken each day Oct-Nov 2012")
```
```{r, echo = TRUE}

```

## What is mean total number of steps taken per day?
#3.Mean and median number of steps taken each day

```{r , echo = TRUE}
TheMean = mean(number_of_steps_per_day$sum)
TheMedian = median(number_of_steps_per_day$sum)
print(TheMean)
print(TheMedian)
```


## What is the average daily activity pattern?

#4.Time series plot of the average number of steps taken

```{r, echo = TRUE}
library(ggplot2)
number_of_steps_per_interval<-ddply(activity_moni_data,.(interval),summarise,sum = sum(steps,na.rm = TRUE))
head(number_of_steps_per_interval)

```
```{r, echo = TRUE}
g<-ggplot(number_of_steps_per_interval,aes(x=interval,y = sum,group =1))
g+geom_line()+ 
xlab(" Interval") + 
ylab("Average Number of Steps") +
ggtitle("Time series plot of the average number of steps taken")
```
```{r, echo = TRUE}

```
#5.The 5-minute interval that, on average, contains the maximum number of steps
```{r, echo = TRUE}
number_of_steps_per_interval[which(number_of_steps_per_interval$sum==(max(number_of_steps_per_interval$sum))),]

```

## Imputing missing values
#6.Code to describe and show a strategy for imputing missing data
```{r, echo = TRUE}

```
```{r, echo = TRUE}

```
Lets count the misssing values

```{r, echo = TRUE}
number_missing_values<-sum(is.na(activity_moni_data))
number_missing_values
```
The number of missing values in the dataset is equal to the number of missing value in the steps which is 2304.
The technique is to replace the missing values by the mean
```{r , echo = TRUE}
na_position<-which(is.na(activity_moni_data$steps))
mean_vec <- rep(mean(activity_moni_data$steps, na.rm=TRUE), times=length(na_position))



```
```{r,echo = TRUE}
activity_moni_data[na_position,"steps"]<-mean_vec
head(activity_moni_data)
```
#7.Histogram of the total number of steps taken each day after missing values are imputed

```{r, echo = TRUE}
Compl_activity_moni_data<-ddply(activity_moni_data,.(date),summarise,sum = sum(steps))
hist(Compl_activity_moni_data$sum,col = "black", main = "Total number of steps taken each day after missing values are imputed")
```
```{r,echo = TRUE}
```


## Are there differences in activity patterns between weekdays and weekends?
#8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
We need to transform the data in order to integrate the weekdays and weekends 
```{r,echo = TRUE}
act1<-read.csv("activity.csv")
act1$date<-as.POSIXct(act1$date,format = "%Y-%m-%d")
```
Let's compute the weekday from the date attributte
```{r,echo =TRUE}
act1<-data.frame(date = act1$date,
                 weekday =  tolower(weekdays(act1$date)),
                 steps = act1$steps,
                 interval= act1$interval)
head(act1)
```
Now lets identify whether the day is weekday or weekend

```{r , echo =TRUE}
act1<-cbind(act1,datetype=ifelse(act1$weekday=="saturday" | act1$weekday == "sunday","weekend","weekday"))
head(act1)
```

```{r , echo =TRUE}
act1<-data.frame(date=act1$date,
                     weekday = act1$weekday, 
                     datetype=act1$datetype,
                     steps = act1$steps,
                     interval = act1$interval)
head(act1)

```
```{r , echo =TRUE}
na_position<-which(is.na(act1$steps))
mean_vec2 <- rep(mean(act1$steps, na.rm=TRUE), times=length(na_position))
act1[na_position,"steps"]<-mean_vec2
head(act1)
```
```{r , echo =TRUE}
library(lattice)
mean_data<-aggregate(act1$steps, 
                       by=list(act1$datetype, 
                               act1$weekday,
                               act1$interval),FUN= mean)
names(mean_data)<-c("datetype", "weekday", "interval", "mean")
head(mean_data)
```
```{r , echo =TRUE}
library(lattice)
xyplot(mean ~ interval|datetype,mean_data,type ="l",lwd=1, xlab="Interval",ylab="Number of steps", layout=c(1,2))

```

